Automatic extraction of symbolic network signatures from malware code

### 2016-11-21 paper.tex

\section{Project Proposal}
\textbf{Tasks}:
\begin{itemize}
	\item Identify and characterize features of common malware code - 1-2 weeks.
	Refer to \Cref{sec:malwaremodel}
	\item \added{Extracting a network signature of length $n$ 
		from the program \\
	Refer to \Cref{sec:extraction}
	\item  Generalizing the network signature extraction and further optimizations
	and evaluation.}
	\item \deleted{Modifying/Extending existing tools like Codesurfer for symbolic 
		network signature extraction - 2-3 weeks}
\end{itemize}

\noindent\textbf{Open source tools}: Static analysis tools, malware repositories

@@

@@

\subsection{Example}
Let us consider the following program receiving two-bit packets:
\begin{lstlisting}	
s1 = 0;
s2 = 0;
while(True) {
	packet = socket.receive();
	if (s1 == 0 and packet[0] == 0) {
	s1 = 1;
	}
	if (s1 == 1 and s2 == 0 and packet[1] == 1) {
	s2 = (not s2) and packet[1];
	}
	if (s1 == 1 and s2 == 1 and packet[0] == 1 and packet[1] == 1) {
	attack()  
	}	
}
\end{lstlisting}	
We denote the state variables by $s_1, s_2$ and packet 
variables by $p_1, p_2$.
For the first \texttt{if} block, 
$\gamma: \neg s_1 \wedge \neg p_1$, and $\delta: s_1'$ ($s_1$ 
is set by the action). Similarily,
for the second block: 
$\gamma: s_1 \wedge \neg s_2 \wedge p_2$, and $\delta: 
s_2' \Leftrightarrow \neg s_2 \wedge p_1$.
For this program, initial state $S_0 = (0, 0)$.
The attack point predicate $\gamma_M$ is as follows:
\begin{equation}
	s_1 \wedge s_2 \wedge p_1 \wedge p_1
\end{equation}
Therefore, if we add the constraints as described in the 
earlier section, for $n=3$, we would obtain a
valid network signature. 
\bibliography{references}

>>>>

\section{Project Proposal}
\textbf{Tasks}:
\begin{itemize}
	\item Identify and characterize features of common malware code - 1-2 weeks
	\added{Refer to sections A and B}
	\item \deleted{Develop the abstract semantics and operators for extracting  
	symbolic network signatures from malware code - 2-3 weeks} \\
	\added{Automatically generate a symbolic automaton of the given program
		based on packet and variable state - 2-3 weeks}
	\item Modifying/Extending existing tools like Codesurfer for symbolic 
	network signature extraction - 2-3 weeks
\end{itemize}

\noindent\textbf{Open source tools}: SMT solvers, malware repositories

@@

\section{Related Work}
An important assumption that we make is that of the simplification of the botnet client code to a set of \textbf {if (state, packet)} statements. Looking at existing research, significant work has gone into automatic protocol reverse-engineering of botnet C\&C protocols using dynamic binary analysis based on network traces. This generally follows a two step process: Extract the protocol grammar, that captures the structure of messages comprising the protocol. Knowing the protocol message format would enable the construction of the protocol state machine, which represents the sequence of messages as a specification of protocol states and transition between states (based on messages received). 
For example, Polyglot~\cite{polyglot} is a system that uses dynamic binary analysis to automatically extract the protocol message format, and Dispatcher~\cite{dispatcher} is a tool that makes use of techniques specified in Polyglot to extract the message format and field semantics for messages sent and received by the application.

Taint propagation is used by the tool to map how the packets received manipulate the state. When the program calls the \textit{read()} function, Dispatcher uses the function's prototype to determine if any argument is tainted based on the memory locations it accesses. Via this approach, a spam botnet called MegaD is analyzed by Dispatcher.

Thus, having knowledge of the network packets received would enable us to make inferences on how they could transition to an 'attack point'. Our work focuses on static analysis of the malware code, without having to execute it and/or capture network traces.

In addition, work has been done on inferring state machines automatically for stateful network protocols - Prospex~\cite{prospex}. Prospex essentially performs clustering of similar network messages and tracks protocol state changes. By merging similar states, a minimal state machine is produced, and the paper produces a labeled state tree for the Agobot malware. This could be used as our initial malware program model for which we determine the sequence of packets to initiate an attack.

@@

\subsection{Example}
Let us consider the following program receiving two-bit packets:
\begin{lstlisting}	
s1 = 0;
s2 = 0;
while(True) {
	packet = socket.receive();
	if (s1 == 0 and packet[0] == 0) {
	s1 = 1;
	}
	if (s1 == 1 and s2 == 0 and packet[1] == 1) {
	s2 = (not s2) and packet[1];
	}
	if (s1 == 1 and s2 == 1 and packet[0] == 1 and packet[1] == 1) {
	attack()  
	}	
}
\end{lstlisting}	
We denote the state variables by $s_1, s_2$ and packet 
variables by $p_1, p_2$.
For the first \texttt{if} block, 
$\gamma: \neg s_1 \wedge \neg p_1$, and $\delta: s_1'$ ($s_1$ 
is set by the action). Similarily,
for the second block: 
$\gamma: s_1 \wedge \neg s_2 \wedge p_2$, and $\delta: 
s_2' \Leftrightarrow \neg s_2 \wedge p_1$.
For this program, initial state $S_0 = (0, 0)$.
The attack point predicate $\gamma_M$ is as follows:
\begin{equation}
	s_1 \wedge s_2 \wedge p_1 \wedge p_1
\end{equation}
Therefore, if we add the constraints as described in the 
earlier section, for $n=3$, we would obtain a
valid network signature.

\bibliographystyle{abbrv}
\bibliography{references}
